---
title: "simpleRegress"
author: "Onalee Garusinghe"
date: "Fall 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
R-squared of simp is 0.87.

Best to Worst:
1. Neck           ARsqrd: 0.87     RST: 43.91


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
agem <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(agem)
summary.lm(agem)
```
R-squared of agemodel is 0.5526

Best to Worst:
1. Chest          ARsqrd: 0.55     RST: 81.47
2. Neck           ARsqrd: 0.87     RST: 43.91


```{r}
headlenmodel <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(headlenmodel)
summary.lm(headlenmodel)
```

R-squred for headlenmodel is 0.69.

Best to Worst:
1. Chest          ARsqrd: 0.92     RST: 33.08
2. Neck           ARsqrd: 0.87     RST: 43.91
3. Head Length    ARsqrd: 0.69     RST: 67.81

```{r}
headwthmodel <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(headwthmodel)
summary.lm(headwthmodel)
```
Headwth model's r-squared is 0.6064.

Best to Worst:
1. Chest          ARsqrd: 0.92     RST: 33.08
2. Neck           ARsqrd: 0.87     RST: 43.91
3. Head Length    ARsqrd: 0.69     RST: 67.81
4. Head Width     ARsqrd: 0.60     RST: 76.41

```{r}
lengthmodel <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(lengthmodel)
summary.lm(lengthmodel)
```

Lengthmodel's r-squared is 0.7423.

Best to Worst:
1. Chest          ARsqrd: 0.92     RST: 33.08
2. Neck           ARsqrd: 0.87     RST: 43.91
3. Length         ARsqrd: 0.74     RST: 61.83
4. Head Length    ARsqrd: 0.69     RST: 67.81
5. Head Width     ARsqrd: 0.60     RST: 76.41

```{r}
agemodel <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(agemodel)
summary.lm(agemodel)
```
R-squared of agemodel is 0.5526

Best to Worst:
1. Chest          ARsqrd: 0.92     RST: 33.08
2. Neck           ARsqrd: 0.87     RST: 43.91
3. Length         ARsqrd: 0.74     RST: 61.83
4. Head Length    ARsqrd: 0.69     RST: 67.81
5. Head Width     ARsqrd: 0.60     RST: 76.41
6. Age            ARsqrd: 0.55     RST: 81.47